|

### 2020-10-14 manuscript/article.Rmd
small edits to article, add to-do to README

---
title:  |
        | Financial Incentives and Health Care Prevention: Evidence from an Experimental *Aedes aegypti* Control Programme in Brazil
author: 
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "We thank Louise Araújo, Livio DiLonardo, Saad Gulzar, Cathy Hafer, Carlo Holz, Dimitri Landa, Guisela Pereira, Adam Przeworski, Catarina Roman, David Skarbek, Denis Stukal, Marcos Tourinho, Scott Tyson, and participants at MPSA, NYU PE Seminar, and SPSA for their valuable comments. Special thanks to XXXX for their excellent research assistance. Replication materials are available at <http://github.com/danilofreire/incentives-health-care>."
abstract:    |
    | __Objectives.__ To quantify the impact of monetary incentives on the performance of health care workers assigned to exterminate *Aedes aegypti*, as well as their effect on dengue hospitalisations. 
    | __Methods.__ We hired and trained subjects to visit households, find *A. aegypti* breeding sites, and kill mosquito larvae in the city of Rio Verde, Brazil, in May 2018. We provided individual and team bonuses to the two treatment groups, while the control group received a flat compensation for completing their tasks.
    | __Results.__ Financial bonuses increased the number of cleaned breeding sites in both treatment status (individual and team bonuses) and team bonuses also improved larvae extermination. The intervention decreased the disease incidence in 10.3%, but the findings are not statistically significant across all model specifications.
    | __Conclusions.__ Performance bonuses improve productivity in public service provision, yet more research is needed to asssess whether higher health care worker performance decreases dengue hospitalisations.
    | __Public Health Implications.__ *A. aegypti* control programmes may benefit from alternative compensation schemes, specially when provided to teams. For this strategy to succeed, financial incentives have to be distributed widely as their aggregate effect is limited.
abstractspacing: single
keywords: Aedes aegypti, dengue, financial incentives, mosquito control, randomised field experiment, preventive health care
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

The *Aedes aegypti* mosquito is the major worldwide vector of several arboviral
diseases, including dengue, yellow fever, Zika, and chikungunya
[@burt2012chikungunya; @jansen2010dengue; @leta2018global;
@thangamani2016vertical]. Over the past 30 years, arboviral incidence has
increased exponentially, with estimates suggesting that dengue is endemic in
around 100 countries and causes 400 million infections and 22 thousands deaths
each year [@bhatt2013global; @cdc2020dengue]. Since there are no safe vaccines
for most *Aedes*-transmitted diseases [@sridhar2018effect;
@world2017dengue], we have seen a number of dengue and Zika outbreaks in Asia,
the Caribbean, South America, and the United States [@carlson2016ecological;
@leta2018global]. Chikungunya virus, which was endemic only to Africa, has
already been reported in every region of the world [@grandadam2011chikungunya;
@kraemer2015global; @mayer2017emergence].

@@

Research shows that financial incentives may increase the productivity of
health care workers [@clemens2014physicians; @muthmainnah2017improving;
@ulrich2005does]. Nevertheless, systematic reviews indicate that the available
evidence has serious methodological shortcomings, such as selection bias and
inadequate experimental blinding, what limits the generalisability of current
findings [@bassani2013financial; @flodgren2011overview; @scott2011effect].
Moreover, it remains unclear which rewarding schemes have a stronger effect on
health care worker performance [@fainman2020design], and whether monetary
incentives have a sustained effect on patient outcomes [@jha2012long;
@langdown2014use].

@@

A typical visit to a household consisted of the following steps. First,
participants rang the doorbell and explained that the municipality had
experienced a dengue fever outbreak. They instructed dwellers on how to lower
dengue incidence and handed them a leaflet with information about dengue
prevention practices. Then, subjects asked whether they were allowed to inspect
house yard. If granted permission, they entered the household and searched for
clean breeding sites, such as pots filled with clear water, and recipients
containing larvae. Upon exterminating a breeding site (pouring out the
accumulated water, for example), subjects had to report their task by taking a
photo on their cellphones. When they discovered larvae, they had to record a
video showing the larvae before exterminating it. At the end of their visit,
they had to take a picture of the household to account for their presence in
the location.

This study evaluates two outcomes. The first outcome is field productivity of
the health care workers. We measure worker productivity by coding cellphone
data collected during the intervention into four indicators: 1) the number of
houses visited; 2) the number of houses visited within less than two minutes;
3) the number of breeding sites removed and cleaned; 4) the number of
exterminated larvae. We analyse the data using ordinary least squares with
robust standard errors [@imbens2016robust]. We estimate the statistical power
of the study assuming a treatment effect of 0.15 standard deviations.

The second outcome of interest is dengue incidence in the municipality of Rio
Verde measured 4, 8, 12, and 16 days after the intervention. The Rio Verde
Mayor's Office provided us with its annual health care report, which contains
the dates and personal information of patients who had dengue fever and
received hospital treatment from January to September 2018. We georeferenced
the home addresses of the patients and compared the incidence map with the
distribution of our intervention. We also compare dengue incidence in each of
the treatment statuses with incidence figures in the control group. The unit of
analysis is the census tract in both models. We employ a
differences-in-differences estimator to assess the treatment effects
[@angrist2008mostly].

@@

>>>>

---
title:  |
        | Financial Incentives and Health Care Prevention: Evidence from an Experimental *Aedes aegypti* Control Programme in Brazil
author: 
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "We thank Louise Araújo, Livio DiLonardo, Saad Gulzar, Cathy Hafer, Carlo Holz, Dimitri Landa, Guisela Pereira, Adam Przeworski, Catarina Roman, David Skarbek, Denis Stukal, Marcos Tourinho, Scott Tyson, and participants at MPSA, NYU PE Seminar, and SPSA for their valuable comments. Special thanks to XXXX for their excellent research assistance. Replication materials are available at <http://github.com/danilofreire/incentives-health-care>."
abstract:    |
    | __Objectives.__ To quantify the impact of financial incentives on the productivity of health personnel combating *Aedes aegypti* and on city-level numbers of dengue hospitalisations.
    | __Methods.__ We hired and trained subjects to visit households, find *A. aegypti* breeding sites, and kill mosquito larvae in the city of Rio Verde, Brazil, in May 2018. We provided individual and team bonuses to two treatment groups, while the control group received a flat compensation for completing their tasks.
    | __Results.__ Financial bonuses increased the number of cleaned breeding sites in both treatment groups (individual and team bonuses) and the collective treatment also improved larvae extermination. The intervention lowered dengue hospitalisations in 10.3%, but the result is not consistent across all model specifications.
    | __Conclusions.__ Performance bonuses improve the performance of health personnel, yet more research is needed to assess whether higher worker productivity decreases dengue hospitalisations.
    | __Public Health Implications.__ *A. aegypti* control programmes may benefit from alternative compensation schemes, specially when provided to teams. For this strategy to succeed, financial incentives have to be distributed widely as their aggregate effect is limited.
abstractspacing: single
keywords: Aedes aegypti, dengue, financial incentives, mosquito control, preventive health care, randomised field experiment, worker productivity
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

@@

The *Aedes aegypti* mosquito is the major worldwide vector of several arboviral
diseases, including dengue, yellow fever, Zika, and chikungunya
[@burt2012chikungunya; @jansen2010dengue; @leta2018global;
@thangamani2016vertical]. Over the past 30 years, arboviral incidence has
increased exponentially, with estimates suggesting that dengue is endemic in
around 100 countries and causes 400 million infections and 22 thousands deaths
each year [@bhatt2013global; @cdc2020dengue]. Since there are no safe vaccines
for most *Aedes*-transmitted diseases [@world2017dengue], we have seen a number
of dengue and Zika outbreaks in Asia, the Caribbean, South America, and the
United States [@carlson2016ecological; @leta2018global]. Chikungunya virus,
previously endemic only to Africa, has already been reported in every region of
the world [@grandadam2011chikungunya; @kraemer2015global; @mayer2017emergence].

@@

Research shows that financial incentives may increase the productivity of
health care workers [@clemens2014physicians; @muthmainnah2017improving;
@ulrich2005does]. Nevertheless, systematic reviews indicate that the
generalisability of the available evidence is limited due to methodological
shortcomings, as several studies suffer from selection bias and inadequate
experimental blinding [@bassani2013financial; @flodgren2011overview;
@scott2011effect]. Moreover, it remains unclear which rewarding schemes have a
stronger effect on health care worker performance [@fainman2020design], and
whether monetary incentives have a sustained effect on patient outcomes
[@jha2012long; @langdown2014use].

@@

A typical visit to a household consisted of the following steps. First,
participants rang the doorbell and explained that the municipality had
experienced a dengue fever outbreak. They instructed dwellers on how to lower
dengue incidence and handed them a leaflet with information about dengue
prevention practices. Then, subjects asked whether they were allowed to inspect
the resident's house yard. If granted permission, they entered the household
and searched for clean breeding sites, such as pots filled with clear water,
and for recipients containing larvae. Upon exterminating a breeding site
(pouring out the accumulated water, for example), subjects had to report their
task by taking a photo on their cellphones. When they discovered larvae,
participants had to record a video showing the larvae before exterminating
them. At the end of their visit, subjects had to take a picture of the household to
account for their presence in the location.

This study evaluates two outcomes. The first outcome is the field productivity
of health care providers. We measure worker productivity by coding cellphone
data collected during the intervention into four indicators: 1) the number of
houses visited; 2) the number of houses visited within less than two minutes;
3) the number of breeding sites removed and cleaned; 4) the number of
exterminated larvae. We analyse the data using ordinary least squares with
robust standard errors [@imbens2016robust]. We estimate the statistical
power of the study assuming a treatment effect of 0.15 standard deviations.

The second outcome of interest is dengue incidence in the municipality of Rio
Verde measured 4, 8, 12, and 16 days after the intervention. The Rio Verde
Mayor's Office provided us with its annual health care report, which contains
the dates and personal information of patients who had dengue fever and
received hospital treatment from January to September 2018. We georeferenced
the home addresses of the patients and compared the incidence map with the
distribution of our intervention. We also compare dengue incidence in each of
the treatment statuses with incidence figures in the control group. In both
models, the unit of analysis is the census tract. We employ a
differences-in-differences estimator to measure the treatment effects
[@angrist2008mostly].

@@

\vspace{.3cm}
